MSE and dynamic state variable model with sigma function

### 2019-04-04 doc/MSE_MAR_18.tex
small text fixes

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Jan Jaap Poos}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{janjaap.poos@wur.nl}}

%Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY). 
\abstract{The EU common fisheries policy (CFP), designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the  maximum sustainable yield (MSY) reference points as targets for exploiting commercially important fish stocks. To incentivise individual business to improve selectivity and avoid unwanted catches, it also introduced a landing obligation. Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests. The impact of exploiting at MSY and no discarding allowance on the optimal effort allocation between fleets is analyzed. We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model. The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations. Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation. Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.4 times larger profits than before the CFP implementation. The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}

@@

where $R(p, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependent over fish age (subsection size dependent pricing). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices. The term $E_{p}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

@@

Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species when species differ only in spatial distribution. Both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat. The changes in production potential were caused by changing exploitation patterns in the fishery due to quota constraints in one species and the resulting spatial reallocation of fishing effort towards the nursery habitat of the other. The resulting change in age selectivity of the fishery decreased the production of the unmanaged stock (Fig. \ref{f:selectivity}).

Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged. However, the stochastic nature of the catches combined with the individual quota result in individual fishers stopping when one of their quotas is reached. At the population level this results in overal cacthes that are lower than the TACs set based on the MSY principle. Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere (xxxx,xxxx,xx), but thereit was attributed to a difference in the life-history characteristics of the different species. Here it is only because of schochasticiyty in catches that is pervasive in all fisheries.DEPENDS ON QUOTA TRADING, AND IF THAT IS POSSIBLE/ HOW GOOD IS MARKET

@@

Moving from an unmanaged period to a situation where one or both species were quota regulated by implementing MSY objectives while maintaining a LO management regime (i.e. main CFP objectives: exploiting at MSY and no discarding allowance) led to a very restrictive fishing activity (Fig. \ref{f:effort}). Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term. In addition to reconciling fishery and conservation objectives (such as the CFP), in agreement with \cite{Grafton2007}, setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig. \ref{f:effort}). However, as \cite{Prellezo2016a} concluded, since effort was drastically reduced to achieve the objectives of the CFP; the increase in net revenue gains were only experienced by some vessels. Consequently, a significant part of the fleet had to cease its activity (Fig. \ref{f:effort}).

@@

 We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY. However, applying any kind of exemption or flexibility beyond the MSY would be negative \cite{Prellezo2016}. Consequently, a critical feature of management bodies is to ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation. Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}. The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance. With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2015, Batsleer2013}. In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations. However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance. IF YOU WANT TO STUDY THIS, THEN DETAILED PARAMETRIZATION IS NEEDED THAT INCLUDES CASE DEPENDENT LH CHARATRISTICS AND MIGRATIONS
 %Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation. The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis

>>>>

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Adriaan Rijnsdorp}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{adriaan.rijnsdorp@wur.nl}
\and
\name{Jan Jaap Poos}
\address{Wageningen University, The Netherlands.}
\email{janjaap.poos@wur.nl}}
%Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY). 
\abstract{The EU common fisheries policy (CFP), designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the maximum sustainable yield (MSY) reference points as targets for exploiting commercially important fish stocks. To incentivise individual business to improve selectivity and avoid unwanted catches, it also introduced a landing obligation. Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests. The impact of exploiting at MSY and no discarding allowance on the optimal effort allocation between fleets is analyzed. We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model. The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations. Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation. Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.4 times larger profits than before the CFP implementation. The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}

@@

where $R(p, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependent over fish age (subsection size dependent pricing). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while $C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices. The term $E_{p}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

@@

Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species. The only prerequisite for this to happen is a difference in the spatial ontogeny of the species: both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat. Changes in the quota constraints led to a spatial redistribution of the fishery. As a result of this spatial redistribution, the age-dependent exploitation for the species changed. For the unconstrained species, the quota resulted in a shift of the fishing pressure towards the younger ages. This in turn decreased the production of the unmanaged stock, as can be seen from the yield curves (Fig. \ref{f:selectivity}).

Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged. However, individual quota cobined with the stochastic fisheries catches resulted in lower than intended exploitation rates. The TACs were set so that the total TACs corresponded to the harvest rate that resulted in maximum sustainable yields. The TACs were divided in individual quotas for each fishing vessel. The lower than intended harvest rates were caused by the fact that despite their planning, fishers often exhausted one of the quotas before the other. If discarding is not allowed, then individual fishers must then stop fishing. At the population levels of the two species this resulted in overal cacthes that were lower than the TACs. Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere (xxxx,xxxx,xx), but thereit was attributed to a difference in the life-history characteristics of the different species. Here it is only because of schochasticiyty in catches that is pervasive in all fisheries.DEPENDS ON QUOTA TRADING, AND IF THAT IS POSSIBLE/ HOW GOOD IS MARKET

@@

Moving from an unmanaged period to a situation where one or both species were quota regulated by implementing MSY objectives while maintaining a LO management regime (i.e. main CFP objectives: exploiting at MSY and no discarding allowance) led to a very restrictive fishing activity (Fig. \ref{f:effort}). Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term. In addition to reconciling fishery and conservation objectives (such as the CFP), setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig. \ref{f:effort}). However, as \cite{Prellezo2016a} concluded, since effort was drastically reduced to achieve the objectives of the CFP; the increase in net revenue gains were only experienced by some vessels. Consequently, a significant part of the fleet had to cease its activity (Fig. \ref{f:effort}).

@@

 We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY. However, applying any kind of exemption or flexibility beyond the MSY would be negative \cite{Prellezo2016}. Consequently, a critical feature of management bodies is to ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation. Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}. The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance. With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Batsleer2013}. In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations. However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance. IF YOU WANT TO STUDY THIS, THEN DETAILED PARAMETRIZATION IS NEEDED THAT INCLUDES CASE DEPENDENT LH CHARATRISTICS AND MIGRATIONS
 %Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation. The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis
